---
title: "02_Map_DM_lifespans"
author: "Hannah Koenker"
date: "11/30/2021"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(maptools)
library(maps)
library(broom)
library(readxl)
library(janitor)
library(spData)
library(sf)
```

First I gather the data - found both these bits of code via search, I'm not exactly sure what the st_transform does but that's ok for now.


``` {r wrld_simpl}
## This file does not have South Sudan:

# data(wrld_simpl)
# afr=wrld_simpl[wrld_simpl$REGION==2,]
# plot(afr)
```

This map source from World Bank has South Sudan, which is better.
```{r worldmap}
africa = world %>% 
  filter(continent == "Africa", !is.na(iso_a2)) %>% 
  left_join(worldbank_df, by = "iso_a2") %>% 
  dplyr::select(name, subregion) %>% 
  st_transform("+proj=aea +lat_1=20 +lat_2=-23 +lat_0=0 +lon_0=25")
plot(africa)

```

```{r dontneed}
# library(raster)
# getData("GADM", country = "SS, level = 2") # get south sudan map
```

I read in my data about how long insecticide-treated nets are lasting in sub-Saharan Africa. They are supposed to last 3 years but frequently they wear out / are thrown away before that.


```{r read}
surv <- read_excel("Median lifespan DM multisite.xlsx") %>% clean_names() %>% 
  separate(gps, into = c("Lati", "Longi"), sep = ",", convert = TRUE) 

# tidy extracts the named columns from the shapefile. Broom works on the SP object.
# afr_surv_fortified <- tidy(afr, region = "NAME")

# make breaks in medianlifespan and convert to factor, drop extraneous variables
surv <- surv %>%
  mutate(lifespan = cut(medianlifespan, breaks = c(0, 1, 2, 3, 4, 5, 6))) %>%
   mutate(lifespan = as.factor(lifespan)) %>% 
  dplyr::select(country, end, iso, district, medianlifespan, bndsurv, lifespan, Lati, Longi)

# note to self: early studies are set as 2.9 median lifespan per Tom Smith's unpublished paper from July 2021, across all brands and sites.

  
```



```{r make_sf}
crs_africa <-  st_crs(africa)

surv_sf <- surv %>% 
  filter(!is.na(Longi)) %>% 
  st_as_sf(coords = c("Longi", "Lati"), crs = 4326) %>% 
  st_transform(crs = crs_africa)

```

``` {r Map2}
ggplot() +
  geom_sf(data = africa,
          fill = "WhiteSmoke",
          color = "white",
          size = 0.5) +
  geom_sf(data = surv_sf,
    aes(fill = lifespan),
    size = 2,
    shape = 21,
    colour = "darkgrey",
    alpha = 8 / 10
  ) +
  theme_void() +
  scale_fill_brewer(
    palette = "Spectral",
    na.value = "WhiteSmoke",
    labels = c(
      "0-1",
      "1-2",
      "2-3",
      "3-4",
      "4-5",
      "5-6",
      "results pending or unpublished"
    )
  ) +
  coord_sf() +
  labs(fill = "Median lifespan (years)",
       title = "Estimated median ITN lifespan reported from field durability studies, 2008-present") +
  theme(plot.margin = grid::unit(c(0, 0, 0, 0), "null"))

ggsave("dm_map_survival2.png")
 
```

``` {r Map3}
# africa_rd <- africa %>% 
#   mutate(rd = case_when(
#     name == "Nigeria" ~ 1,
#     name == "Tanzania" ~ 1,
#     name == "DRC" ~ 1,
#     name == "Mozambique" ~ 1,
#     TRUE ~ 2))    
# 
# cty_fills <- c("WhiteSmoke", "Teal")
# 
#   ggplot() +
#   geom_sf(data = africa_rd,
#           fill = rd,
#           color = "white",
#           size = 0.5) 
#   
#   
#   +
#   geom_sf(data = surv_sf,
#     aes(fill = lifespan),
#     size = 2,
#     shape = 21,
#     colour = "darkgrey",
#     alpha = 8 / 10
#   ) +
#   theme_void() +
#   scale_fill_brewer(
#     palette = "Spectral",
#     na.value = "WhiteSmoke",
#     labels = c(
#       "0-1",
#       "1-2",
#       "2-3",
#       "3-4",
#       "4-5",
#       "5-6",
#       "results pending or unpublished"
#     )
#   ) +
#   coord_sf() +
#   labs(fill = "XXX",
#        title = "Sites included in RD Score Analysis") +
#   theme(plot.margin = grid::unit(c(0, 0, 0, 0), "null"))
# 
# ggsave("dm_map_RD_sites.png")
 
```


```{r Map4}

rd <- surv_sf %>% 
  filter(end == 2018 | end == 2019,
         !is.na(lifespan),
         country != "Ethiopia")

ggplot() +
  geom_sf(data = africa,
          fill = "WhiteSmoke",
          color = "white",
          size = 0.5) +
  geom_sf(data = rd,
    aes(fill = lifespan),
    size = 4,
    shape = 21,
    colour = "darkgrey",
    alpha = 8 / 10
  ) +
  theme_void() +
  scale_fill_brewer(
    palette = "Spectral",
    na.value = "WhiteSmoke",
    labels = c(
      "0-1",
      "1-2",
      "2-3",
      "3-4",
      "4-5",
      "5-6",
      "results pending or unpublished"
    )
  ) +
  coord_sf() +
  labs(fill = "Median lifespan (years)",
       title = "Sites included in RD Analysis") +
  theme(plot.margin = grid::unit(c(0, 0, 0, 0), "null"))

ggsave("dm_map_rd2.png")
```

